---
title: "R Notebook"
output: html_notebook
---

```{r}
library(here)
library(tidyverse)
library(ggpubr)
library(epitools)
pal2 <- c("#015b58", "#5962b5")
pal2.flip <- c("#5962b5", "#015b58")
here::here()



box <- read.delim2(here("data/inphared_db/14Apr2025inph_0A_v3.txt" ))
box <- subset(box, box$Header!="Header")
box <- subset(box, box$Partial=="None")
box <- box %>% rename(product = Sequence, phage = Contig)
box[,c(2)] <- "0A_box"
#box <- box[,c(3,2,1)]
box$count <- 1


gc_content <- function(seq) {
  seq <- toupper(seq)
  bases <- strsplit(seq, "")[[1]]
  gc_count <- sum(bases %in% c("G", "C"))
  total <- length(bases)
  return(gc_count / total)
}

# Vectorized version for multiple sequences
gc_content_vec <- function(seqs) {
  sapply(seqs, gc_content)
}

box.f <- box

box.f$GC.flanks <- gc_content_vec(box.f$Context_no_motif)
box.f$GC.Up <- gc_content_vec(box.f$Upstream)
box.f$GC.Down <- gc_content_vec(box.f$Downstream)


```
### DECIDE CUT OFF FOR AT RICHNESS

```{r}

## no filter
#box <- box.f


## gc filtering
#box <- subset(box.f, box.f$GC.flanks<0.35)
box <- subset(box.f, box.f$GC.Up<0.35)
#box <- subset(box.f, box.f$GC.Down<0.35)

```



```{r}
box <-  box %>% group_by(phage, product) %>% summarise(sum=sum(count)) 

phage <- read.csv(here("data/inphared_db/14Apr2025_knownsporestatus.csv"), row.names=1)

all <- merge(box, phage, by.x="phage", by.y="Accession", all.x=TRUE, all.y=TRUE) 

all$product[is.na(all$product)] <- "0A_box"

all$sum[is.na(all$sum)] <- 0

all[is.na(all)] <- "currently_missing"
all <- subset(all, all$newgtdb_Phylum!="currently_missing")

all$HostPhyla <- "OtherPhyla"
all$HostPhyla <- ifelse(all$newgtdb_Phylum=="Bacillota", "Bacillota", "OtherPhyla")

all$sporulation <- "unknown"
all$sporulation <- ifelse(all$f_spor=="TRUE", "Spor", all$sporulation)
all$sporulation <- ifelse(all$f_spor=="FALSE", "Nonspor", all$sporulation)

all <- subset(all, all$sporulation!="unknown" )
all <- subset(all, all$predicted_label!="missing")

all <- unite(all, "phyla_spor_type", c("HostPhyla", "sporulation", "predicted_label"), sep = "_", remove = FALSE, na.rm = FALSE)

all <- unite(all, "spor_type", c("sporulation", "predicted_label"), sep = "_", remove = FALSE, na.rm = FALSE)

all <- unite(all, "specphyla_spor_type", c("newgtdb_Phylum", "sporulation", "predicted_label"), sep = "_", remove = FALSE, na.rm = FALSE)

all$genome <- as.numeric(all$Genome.Length..bp)
all$phageGC <- as.numeric(all$molGC...)


all$genome <- log10(all$genome)

```



#### quick sanity check about phage GC 


```{r}

ggplot(all, aes(x = factor(host_phage_spor), fill = factor(host_phage_spor), y = phageGC )) +
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2) + ylab("Phage Genome GC%")

ggsave("flanking_qc/phagegc_bytype.png")


 ggplot(all, aes(x = factor(host_phage_spor), y = phageGC)) +
  geom_boxplot(outlier.shape = NA) +  # Remove default outliers so jitter isn't obscured
  geom_jitter(aes(color = sum), width = 0.2, size = 2, alpha = 0.5) +
  scale_color_gradient(low = "lightblue", high = "darkred") +  # Customize gradient as needed
  ylab("GC % of Phage Genome") +
  xlab("Host sporulation") + ggtitle("At least 65%  AT Upstream")
  theme_minimal()
```


phage.gc.flank
ggsave("flanking_qc/phageGC_fullflank.png")
phage.gc.up 
ggsave("flanking_qc/phageGC_upstream.png")
phage.gc.down
ggsave("flanking_qc/phageGC_downstream.png")
phage.gc.unfilt
ggsave("flanking_qc/phageGC_unfilt.png")





```{r}
all$tally <- 1
all$hit <- ifelse(all$sum!=0, 1, 0)

counts <- all %>% group_by(phyla_spor_type) %>% summarise(num.phage=sum(tally), total.0A=sum(sum), mean.0A=mean(sum), hits.0A=sum(hit)) 


### only accounting for sporulation and lifestyle
ggplot(all, aes(x = factor(spor_type), fill = factor(spor_type), y = sum )) +
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2) + ylab("Number of 0A boxes")


## accounting for bacilliota vs. other phyla
ggplot(all, aes(x = factor(phyla_spor_type), fill = factor(phyla_spor_type), y = sum )) +
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2) + ylab("Number of 0A boxes") + ggtitle("# of 0A boxes with <35% upstream flanking GC% ")
```
no.filt
ggsave("flanking_qc/phage0A_sum_nofilt.png")
flank.both
ggsave("flanking_qc/phage0A_sum_flankboth.png")
up.flank
ggsave("flanking_qc/phage0A_sum_flankupstream.png")
down.flank
ggsave("flanking_qc/phage0A_sum_flankdown.png")



```{r}
## accounting for bacilliota vs. other phyla
ggplot(all, aes(x = factor(newgtdb_Phylum), fill = factor(predicted_label), y = sum )) +
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2) + ylab("Number of 0A boxes")

ggplot(all, aes(x = factor(specphyla_spor_type), fill = factor(predicted_label), y = sum )) +
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2) + ylab("Number of 0A boxes")



ggplot(all, aes(x = factor(specphyla_spor_type), fill = factor(specphyla_spor_type), y = sum )) +
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2) + ylab("Number of 0A boxes")



ggplot(all, aes(genome, sum, fill=host_phage_spor, colour = host_phage_spor)) +
  geom_point() +
   geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0,

ggplot(all, aes(phageGC, sum, fill=host_phage_spor, colour = host_phage_spor)) +
  geom_point() +
   geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )
```

```{r}
a1 <- aov(sum ~ phyla_spor_type, data = all)
summary(a1)

TukeyHSD(a1)
```

```{r}
all.spor.counts  <- all %>% group_by(sporulation) %>% summarise(num.phage=sum(tally), total.0A=sum(sum), mean.0A=mean(sum), hits.0A=sum(hit)) 

all.spor.counts$perc.hit <- all.spor.counts$hits.0A/all.spor.counts$num.phage

all.spor.counts
```

```{r}
spor.stat <- c('Spore', 'Non-spore')
outcome <- c('Hits', 'None')
all.hits <- matrix(c(736, 75, 18825, 887), nrow=2, ncol=2, byrow=TRUE)
dimnames(all.hits) <- list('Spore'=spor.stat, 'Outcome'=outcome)
all.hits


fisher.test(all.hits)  

oddsratio(all.hits)

```

other phyla lifestyle 
```{r}

phyla <- subset(all, all$HostPhyla=="OtherPhyla")


phyla.life <- phyla %>% group_by(predicted_label) %>% summarise(num.phage=sum(tally), total.0A=sum(sum), mean.0A=mean(sum), hits.0A=sum(hit)) 

phyla.life

```

```{r}
phyla.stat <- c('Lytic', 'Temp')
outcome <- c('Hits', 'None')
phyla.hits <- matrix(c(7047, 239, 9544, 350), nrow=2, ncol=2, byrow=TRUE)
dimnames(phyla.hits) <- list('Lifestyle'=phyla.stat, 'Outcome'=outcome)
phyla.hits


fisher.test(phyla.hits)  

oddsratio(phyla.hits)

```

##### BACILLIOTA ONLY 

```{r}

baci <- subset(all, all$HostPhyla=="Bacillota")

ggplot(baci, aes(x = factor(spor_type), fill = factor(spor_type), y = sum )) +
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2) + ylab("Number of 0A boxes")



ggplot(baci, aes(genome, sum, fill=spor_type, colour = spor_type)) +
  geom_point() +
   geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0,

```


```{r}
spor.counts <- baci %>% group_by(sporulation) %>% summarise(num.phage=sum(tally), total.0A=sum(sum), mean.0A=mean(sum), hits.0A=sum(hit)) 

spor.counts$perc.hit <- spor.counts$hits.0A/spor.counts$num.phage

spor.counts

```

```{r}
spor.stat <- c('Spore', 'Non-spore')
outcome <- c('Hits', 'None')
bacil.hits <- matrix(c(736, 75, 2234, 298), nrow=2, ncol=2, byrow=TRUE)
dimnames(bacil.hits) <- list('Spore'=spor.stat, 'Outcome'=outcome)
bacil.hits
library(epitools)

fisher.test(bacil.hits)  
oddsratio(bacil.hits)

```


```{r}

virbac <- subset(baci, baci$predicted_label=="virulent")

life.counts <- virbac %>% group_by(sporulation) %>% summarise(num.phage=sum(tally), total.0A=sum(sum), mean.0A=mean(sum), hits.0A=sum(hit)) 

life.counts$perc.hit <- life.counts$hits.0A/life.counts$num.phage

life.counts
```


```{r}
spor.stat <- c('Spore', 'Non-spore')
outcome <- c('Hits', 'None')
vir.hits <- matrix(c(338, 19, 1097, 246), nrow=2, ncol=2, byrow=TRUE)
dimnames(vir.hits) <- list('Spore'=spor.stat, 'Outcome'=outcome)

vir.hits

library(epitools)
fisher.test(vir.hits)  
oddsratio(vir.hits)

```



```{r}

tempbac <- subset(baci, baci$predicted_label=="temperate")

tlife.counts <- tempbac %>% group_by(sporulation) %>% summarise(num.phage=sum(tally), total.0A=sum(sum), mean.0A=mean(sum), hits.0A=sum(hit)) 

tlife.counts$perc.hit <- tlife.counts$hits.0A/tlife.counts$num.phage

tlife.counts
```


```{r}
spor.stat <- c('Spore', 'Non-spore')
outcome <- c('Hits', 'None')
temp.hits <- matrix(c(398, 56, 1383, 52), nrow=2, ncol=2, byrow=TRUE)
dimnames(temp.hits) <- list('Spore'=spor.stat, 'Outcome'=outcome)

temp.hits

library(epitools)
fisher.test(temp.hits)  
oddsratio(temp.hits)

```