---
title: "R Notebook"
output: html_notebook
---

## INPHARED descriptive stats


```{r}
library(tidyverse)
library(here)
library(cowplot)
here()
dr_here()

pal2 <- c("#015b58", "#5962b5")
pal2.flip <- c("#5962b5", "#015b58")
#inph <- read.csv(here("data/inphared_db/14Apr2025_knownsporestatus.csv", row.names=1))
inph <- read.csv(here("00_data", "inphared_db", "14Apr2025_knownsporestatus.csv"), row.names = 1)

inph$sporulation <- "T"
inph$sporulation <- ifelse(inph$f_spor==TRUE, "SproulatingHost", "NonsporulatingHost")

inph <- unite(inph, "phage_type", c("sporulation", "predicted_label"), sep = "_", remove = FALSE, na.rm = FALSE)



### remove duplicates (keep only reference strains)
### currently this only works on bacilliota, 99.5% ANI over 100% alignment threshold
## rest of inphared is pending 

inph <- subset(inph, inph$ref=="not_run" | inph$ref=="ref_seq")

```

### compare genome size of temperate vs. lytic  BACILLOTA ONLY
```{r}
inph.all <- inph
#inph<- subset(inph.all, inph.all$Jumbophage==FALSE)

inph$genome.log <- log10(inph$Genome.Length..bp.)

inph.subset <- subset(inph.all, inph.all$newgtdb_Phylum=="Actinomycetota" | inph.all$newgtdb_Phylum=="Pseudomonadota" | inph.all$newgtdb_Phylum=="Bacteroidota" | inph.all$newgtdb_Phylum=="Bacillota" | inph.all$newgtdb_Phylum=="Campylobacterota" | inph.all$newgtdb_Phylum== "Cyanobacteriota")


inph.bac <- subset(inph.all, inph.all$newgtdb_Phylum=="Bacillota")


inph.log <- ggplot(inph.bac, aes(x = factor(lifestyle), fill = factor(lifestyle), y = genome.log)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("log10 (Genome Size)") + xlab("") + ggtitle("INPHARED") + labs(fill="Lifestyle") + facet_wrap(~ newgtdb_Phylum2, ncol = 2)

inph.genes <- ggplot(inph.bac, aes(x = factor(lifestyle), fill = factor(lifestyle), y = Number.CDS)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Number of Genes") + xlab("") + ggtitle("INPHARED") + labs(fill="Lifestyle") + facet_wrap(~ newgtdb_Phylum2, ncol = 2)

ggplot(inph.subset, aes(x = factor(lifestyle), fill = factor(lifestyle), y = genome.log)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("log10 (Genome Size)") + xlab("") + ggtitle("INPHARED") + labs(fill="Lifestyle") + facet_wrap(~ newgtdb_Phylum2, ncol = 3)


oneway.inph.l <- aov(genome.log ~ lifestyle , data = inph.bac)
oneway.inph.s <- aov(genome.log ~ sporulation , data = inph.bac)

twoway.inph.a <- aov(genome.log ~ lifestyle + sporulation, data = inph.bac)

twoway.inph.m <- aov(Number.CDS ~ lifestyle * sporulation, data = inph.bac)

summary(twoway.inph)

em.inph <- emmeans(twoway.inph.m, ~ lifestyle * sporulation, data=inph.bac)
pairs(em.inph)
summary(em.inph)
contrast(em.inph, method = "pairwise", by = "lifestyle", adjust = "holm")


library(AICcmodavg)

model.set <- list(oneway.inph.l, oneway.inph.s, twoway.inph.a, twoway.inph.m)
model.names <- c("lifestyle", "sporulation", "two.way", "interaction")

aictab(model.set, modnames = model.names)

ggplot(inph.bac, aes(x = factor(lifestyle), fill = factor(lifestyle), y = molGC....)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Phage Genome Size (bp)") + xlab("") + ggtitle("INPHARED, with jumbophages") + labs(fill="Lifestyle") + facet_wrap(~ newgtdb_Phylum2, ncol = 2)



ggplot(inph.bac, aes(x = factor(lifestyle), fill = factor(lifestyle), y = tRNAs)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Phage Genome Size (bp)") + xlab("") + ggtitle("INPHARED") + labs(fill="Lifestyle") + facet_wrap(~ newgtdb_Phylum2, ncol = 2)



ggplot(inph.bac, aes(x = factor(lifestyle), fill = factor(lifestyle), y = Number.CDS)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Phage Genome Size (bp)") + xlab("") + ggtitle("INPHARED, with jumbophages") + labs(fill="Lifestyle") + facet_wrap(~ newgtdb_Phylum2, ncol = 2)

````





ggplot(inph.all, aes(x = factor(f_spor), fill = factor(lifestyle), y = Genome.Length..bp.)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Phage Genome Size (bp)") + xlab("") + ggtitle("INPHARED, with jumbophages") + labs(fill="Lifestyle") 


ggplot(inph.all, aes(x = factor(lifestyle), fill = factor(lifestyle), y = Genome.Length..bp.)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Phage Genome Size (bp)") + xlab("") + ggtitle("INPHARED, with jumbophages") + labs(fill="Lifestyle") + facet_wrap(~ newgtdb_Phylum2, ncol = 2)


ggplot(inph.subset, aes(x = factor(lifestyle), fill = factor(lifestyle), y = Genome.Length..bp.)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Phage Genome Size (bp)") + xlab("") + ggtitle("INPHARED, with jumbophages") + labs(fill="Lifestyle") + facet_wrap(~ newgtdb_Phylum2, ncol = 2)


# ggplot(inph, aes(x = factor(f_spor), fill = factor(predicted_label), y = Genome.Length..bp.)) 
#  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("Phage # Genome Size (bp)") + xlab("") + ggtitle("INPHARED, no jumbophages") + labs(fill="Lifestyle")
  
  
  
  
  
inph.spor <- subset(inph.all, inph.all$f_spor==TRUE)  

size.spor <- ggplot(inph.spor, aes(x = factor(lifestyle), fill = factor(lifestyle), y = Genome.Length..bp.)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Phage Genome Size (bp)") + xlab("") + ggtitle("") + labs(fill="Lifestyle")+ ylim(0, 750000)



inph.no <- subset(inph.all, inph.all$f_spor==FALSE)  

size.no <- ggplot(inph.no, aes(x = factor(lifestyle), fill = factor(lifestyle), y = Genome.Length..bp.)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2) + ylab("Phage Genome Size (bp)") + xlab("") + ggtitle("") + labs(fill="Lifestyle") + ylim(0, 750000) + theme(legend.position = "none")





genome.size <- plot_grid(size.no, size.spor, labels = c('Nonsporulating Host', 'Sporulating Host'), label_size = 12, rel_widths = c(1.5, 2))

genome.size


```{r}

inph.all$f_spor <- as.factor(inph.all$f_spor)
inph.all$predicted_label <- as.factor(inph.all$predicted_label)
inph.all$Host <- as.factor(inph.all$Host)


#inph <- subset(inph, inph$Jumbophage==FALSE)

oneway <- aov(Genome.Length..bp. ~ predicted_label, data = inph.all)

summary(oneway)


twoway <- aov(Genome.Length..bp. ~ predicted_label + f_spor, data = inph.all)

summary(twoway)


twoway.int <- aov(Genome.Length..bp. ~ predicted_label * f_spor, data = inph.all)

summary(twoway.int)


block <- aov(Genome.Length..bp. ~predicted_label + f_spor + Host, data = inph.all)

summary(block)

block_int <- aov(Genome.Length..bp. ~ predicted_label + f_spor * Host, data = inph.all)
summary(block_int)
all_int <- aov(Genome.Length..bp. ~ predicted_label * f_spor * Host, data = inph.all)
summary(all_int)

library(AICcmodavg)

model.set <- list(oneway, twoway, twoway.int, block, block_int)
model.names <- c("one.way", "two.way", "interaction", "blocking", "block_int")

aictab(model.set, modnames = model.names)

par(mfrow=c(2,2))
plot(twoway.int)
par(mfrow=c(1,1))

terms(block)

TukeyHSD(block)  # or:
TukeyHSD(block, which = c("predicted_label", "f_spor"))


#tkey

kruskal.test(predicted_label ~ Genome.Length..bp., data = inph)




library(emmeans)
em <- emmeans(block, ~ predicted_label * f_spor)
pairs(em)

## no spor temp vs. vir, different
## spor temp vs. vir, different

## no difference between spor temp and nonspor temp
## no difference between spor vir and nonspor vir


```

```{r}
 ggplot(inph.subset, aes(x = factor(predicted_label), fill = factor(predicted_label), y = molGC....)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("GC %") + xlab("") + ggtitle("") + labs(fill="Lifestyle")+ ylim(0, 75)+ facet_wrap(~ newgtdb_Phylum2, ncol = 2)

gc.spor <- ggplot(inph.spor, aes(x = factor(predicted_label), fill = factor(predicted_label), y = molGC....)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("GC %") + xlab("") + ggtitle("") + labs(fill="Lifestyle")+ ylim(0, 75)



inph.no <- subset(inph.all, inph.all$f_spor==FALSE)  

gc.no <- ggplot(inph.no, aes(x = factor(predicted_label), fill = factor(predicted_label), y = molGC....)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("GC %") + xlab("") + ggtitle("") + labs(fill="Lifestyle") + ylim(0, 75) + theme(legend.position = "none")




gc.amount <- plot_grid(gc.no, gc.spor, labels = c('Nonsporulating Host', 'Sporulating Host'), label_size = 12, rel_widths = c(1.5, 2))

gc.amount
```


```{r}

inph.all$f_spor <- as.factor(inph.all$f_spor)
inph.all$predicted_label <- as.factor(inph.all$predicted_label)
inph.all$Host <- as.factor(inph.all$Host)

#inph <- subset(inph, inph$Jumbophage==FALSE)

oneway <- aov(molGC.... ~ predicted_label, data = inph.all)

summary(oneway)


twoway <- aov(molGC.... ~ predicted_label + f_spor, data = inph.all)

summary(twoway)


twoway.int <- aov(molGC.... ~ predicted_label * f_spor, data = inph.all)

summary(twoway.int)


block <- aov(molGC.... ~predicted_label + f_spor + Host, data = inph.all)

summary(block)

block_int <- aov(molGC.... ~ predicted_label + f_spor * Host, data = inph.all)
summary(block_int)
all_int <- aov(molGC.... ~ predicted_label * f_spor * Host, data = inph.all)
summary(all_int)

library(AICcmodavg)

model.set <- list(oneway, twoway, twoway.int, block, block_int)
model.names <- c("one.way", "two.way", "interaction", "blocking", "block_int")

aictab(model.set, modnames = model.names)
summary(block)

par(mfrow=c(2,2))
plot(twoway.int)
par(mfrow=c(1,1))

terms(block)

TukeyHSD(block)  # or:



#tkey

kruskal.test(predicted_label ~ molGC...., data = inph)




library(emmeans)
em <- emmeans(block, ~ predicted_label * f_spor)
pairs(em)

### no effect 
n

```


```{r}
tRNA.spor <-  ggplot(inph.spor, aes(x = factor(predicted_label), fill = factor(predicted_label), y = tRNAs)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("# of tRNAs") + xlab("") + ggtitle("") + labs(fill="Lifestyle")+ ylim(0, 30)

 ggplot(inph.subset, aes(x = factor(predicted_label), fill = factor(predicted_label), y = tRNAs)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("# of tRNAs") + xlab("") + ggtitle("") + labs(fill="Lifestyle")+ ylim(0, 30) + facet_wrap(~ newgtdb_Phylum2, ncol = 2)

inph.no <- subset(inph.all, inph.all$f_spor==FALSE)  

tRNA.no <-ggplot(inph.no, aes(x = factor(predicted_label), fill = factor(predicted_label), y = tRNAs)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("# of tRNAs") + xlab("") + ggtitle("") + labs(fill="Lifestyle") + ylim(0, 30) + theme(legend.position="none")


tRNA.amount <- plot_grid(tRNA.no, tRNA.spor, labels = c('Nonsporulating Host', 'Sporulating Host'), label_size = 12, rel_widths = c(1.5, 2))

tRNA.amount

twoway_block <- aov(tRNAs ~ predicted_label+ f_spor+Host, data = inph.all)

summary(twoway_block)

library(emmeans)
em <- emmeans(twoway_block, ~ predicted_label * f_spor)
pairs(em)

### lifestyle, sporulation, and host, have impact on tRNAs 
## non spor vir vs. spor temp are different
## spor vir vs. spor temp are diffent

## vir spor vs. non spor are NOT different
## temp spor vs. non spor ARE NOT different

```


```{r}

ggplot(inph.subset, aes(x = factor(predicted_label), fill = factor(predicted_label), y = Number.CDS)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("# of Coding Regions") + xlab("") + ggtitle("") + labs(fill="Lifestyle")+ ylim(0, 350) + facet_wrap(~ newgtdb_Phylum2, ncol = 2)




CDS.spor <- ggplot(inph.spor, aes(x = factor(predicted_label), fill = factor(predicted_label), y = Number.CDS)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("# of Coding Regions") + xlab("") + ggtitle("") + labs(fill="Lifestyle")+ ylim(0, 350)





inph.no <- subset(inph.all, inph.all$f_spor==FALSE)  

CDS.no <- ggplot(inph.no, aes(x = factor(predicted_label), fill = factor(predicted_label), y = Number.CDS)) + 
  geom_boxplot(binaxis = "y", stackdir = "center", position = "dodge") + geom_jitter(width = 0.2)  + scale_fill_manual(values=pal2, labels = c ("Temperate", "Lytic")) + ylab("# of Coding Regions") + xlab("") + ggtitle("") + labs(fill="Lifestyle") + ylim(0, 350) + theme(legend.position="none")


CDS.amount <- plot_grid(CDS.no, CDS.spor, labels = c('Nonsporulating Host', 'Sporulating Host'), label_size = 12, rel_widths = c(1.5, 2))

CDS.amount


twoway <- aov(Number.CDS ~ predicted_label+ f_spor+Host, data = inph.all)
## lifestyle and sporulation both an impact on number of CDS



summary(twoway)

library(emmeans)
em <- emmeans(twoway, ~ predicted_label * f_spor)
pairs(em)

## lifestyle and sporulation both an impact on number of CDS <0.001

## nonspor temp vs. vir are different
## spor temp vs. vir are different

## vir spor vs. vir no spor are NOT different
## temp spor vs. temp no spor are NOT different


```

```{r}
all <- plot_grid(
  genome.size, gc.amount, CDS.amount, tRNA.amount, ncol=1)
 # labels = c('Genome Size', 'GC %', 'Predicted CD Regions', 'Predicted tRNAs'), ncol = 1)

all

##ggsave(here("data/inphared_db/inphared_desc.png"), height=11, width=8)

```

```{r}
#inph.spor 
#inph.no

## BACILLOTA

library(ggpubr)

inph.bac <- subset(inph.all, inph.all$newgtdb_Phylum=="Bacillota")

gvT <- ggplot(inph.bac, aes(Genome.Length..bp., tRNAs, fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() +
  geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0, 400000)

gvT

cvT <- ggplot(inph.bac, aes(Number.CDS, tRNAs, fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() +
  geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0, 400000)

MvT <- ggplot(inph.bac, aes(molGC...., tRNAs, fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() +
  geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0, 400000)

MvT  ## not interesting lol

gvC <- ggplot(inph.bac, aes(Genome.Length..bp., Number.CDS, fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() +
  geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "left", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0, 400000)


GvGC <- ggplot(inph.bac, aes(Genome.Length..bp., molGC...., fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() + geom_smooth(method="lm") +
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )
GvGC

all.line <- plot_grid(
  gvT, cvT, gvC, GvGC, ncol=1)
 # labels = c('Genome Size', 'GC %', 'Predicted CD Regions', 'Predicted tRNAs'), ncol = 1)

all.line

#ggsave(here("data/inphared_db/inphared_linear.png"), height=11, width=8)


```


```{r}


length.by.codon <- ggplot(inph.bac, aes(genome.log, Number.CDS, fill=host_phage_spor, colour = host_phage_spor, shape = host_phage_spor)) +
  geom_point() + geom_smooth(method="lm") +
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )
length.by.codon

ggsave(here("01_desc_stats/inph_linear_bacil.png"))


```

```{r}
#inph.spor 
#inph.no

## all

library(ggpubr)

gvT <- ggplot(inph.all, aes(Genome.Length..bp., tRNAs, fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() +
  geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0, 400000)


cvT <- ggplot(inph.all, aes(Number.CDS, tRNAs, fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() +
  geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0, 400000)

MvT <- ggplot(inph.all, aes(molGC...., tRNAs, fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() +
  geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0, 400000)

MvT  ## not interesting lol

gvC <- ggplot(inph.all, aes(Genome.Length..bp., Number.CDS, fill=phage_type, colour = phage_type, shape = f_spor)) +
  geom_point() +
  geom_smooth(method="lm")+ # Add regression line
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "left", label.y.npc = "top" )# Add equation#+ ylim(0, 25) #+ xlim(0, 400000)


GvGC <- ggplot(inph.bac, aes(Genome.Length..bp., molGC...., fill=host_phage_spor, colour = host_phage_spor, shape = host_phage_spor)) +
  geom_point() + geom_smooth(method="lm") +
 stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")),
                        label.x.npc = "middle", label.y.npc = "top" )
GvGC

all.line <- plot_grid(
  gvT, cvT, gvC, GvGC, ncol=1)
 # labels = c('Genome Size', 'GC %', 'Predicted CD Regions', 'Predicted tRNAs'), ncol = 1)

all.line

#ggsave(here("data/inphared_db/inphared_linear.png"), height=11, width=8)


```

